\protect Connecting Every Bit of Knowledge: \\ Wikipedia's First Link Network Unravelled

### 2015-11-30 writeup/2015-11-wikipedia-network.tex
synonyms figure + edits

\title{\protect
A Web of Human Knowledge: 
Unravelling Wikipedia's First Link Network
}

@@

"Philosophy" and other philosophical concepts with many traversal visits
are not among the highest-ranking articles by indegree.
"Philosophy" has an indegree of 581, with direct first links from articles about Philosophers and areas of Philosophy: "Existentialism and Humanism," "Predeterminism," "Synoptic Philosophy,"Qualia," "Dorothy Emmet," and "Christopher W. Morris."
While many articles accumulate at "Philosophy" (see traversal visits discussion), 
the accumulation is not the 
result of many articles directly referencing "Philosophy." 
Instead, the accumulation of first links, as we argue in our 
discussion of traversal visits, flows towards Philosophy as the 
ultimate anchor when genralizing from specific articles towards broader notions.
\begin{figure*}[tp!]
  \centering	
  \includegraphics[width=\columnwidth]{graphics/ndegree_loglog.png}
  \caption{
    \textbf{FLN Degree Distribution}
  }
  We construct a log-log plot fit our results to a linear model. The result is a 
  an excellent fit with r = -0.98, yielding a power law exponent of -0.79. 
  The distribution appears scale free with most articles having fewer than 9 
  direct first links, while few hold most.
  \label{fig:degree distribution}
\end{figure*}
The FLN's indegree exihibits a scale-free distribution where a few articles 
receive a most direct first references, while most articles receive few or none.
The average indegree for all 11 million articles is 3.6 direct first links with a standard deviation of 89.5 links.
Only 4,826 articles have more than 100 direct first links and $75\%$ of articles
have fewer than 9. 
When fit to a linear model on a log-log scale (log(rank) versus log(indegree)), 
the model's r value is $-0.98$ suggesting a strong log-log linear fit 
with a power law exponent of $-0.79$.

@@

As a distribution, the number traversal visits per article appears scale-free. The majority of articles have fewer than 30 traversal visits, while few 
have 5 order of magnitude more traversal visits. 
Specifically, $99.76\%$ of articles have fewer than $100$ traversal visits; nearly $80\%$ have none. 
Meanwhile, the highest ranking 30 articles have an extremely disproportionate number of traversal visits.

@@

One means to measure the specificity of an article is to identify the number of synonyms available for a word or topic. The reasoning here is that a  broader topic likely has many more synonyms relative to a specific, concrete topic. Banana has fewer synonyms than botanical for example. To quantify the observed flow, we measured the number of synonyms the article title contained in WordNet---the largest lexical database of the English language 
\cite{wordnet}. 
We find the highest ranking 100 articles have on average 5 more synonyms than the typical article; a difference of 2.5 synonyms if we compare the median number of synonyms in each group. 
As suggested by the median, many articles have no synonyms as we might expect, because titles with more than one word are not likely to appear in a thesaurus. 
Since many articles have no synonyms, we also compared the number of synonyms in the highest-ranking versus typical article, this time excluding all articles without at least 1 synonym. 
We still find the highest ranking 100 articles with an average of 9.0 (median of 7.0) synonyms whereas the remaining articles on average 5.8 synonyms (median of 7.0), even with the exclusion of all articles without any synonyms.
The quantifiable difference in synonyms corroborates the flow of links from concrete, specific articles to broader disciplines or fundamental notions.

@@

\begin{figure*}[tp!]
  \centering	
  \includegraphics[width=\textwidth]{graphics/2_cycles.pdf}
  \caption{
    \textbf{highest ranking 2-cycles}
  }
  \label{fig:2-cycles}

@@

\begin{figure*}[tp!]
  \centering	
  \includegraphics[width=\textwidth]{graphics/3_cycles.pdf}
  \caption{
    \textbf{highest ranking 3-cycles}
  }
  \label{fig:3-cycles}
\end{figure*}

@@

Ranking articles by the number of traversal funnels we find 
"Philosophy" as by far the highest-ranking article with 
$7.37$ million paths
(see figure~\ref{fig:Funnels}).
\begin{figure*}[tp!]
  \centering	
  \includegraphics[width=\textwidth]{graphics/funnels.pdf}
  \caption{
    \textbf{Funnels}
  }
  \label{fig:Funnels}
\end{figure*}
Of any article, the number of traversal funnels Philosophy holds exceeds 
all others by at least two orders of magnitude.
The "Philosophy" cycle which contains "Existence," "Awareness," "Reality," 
and similar articles accumulates the overwhelming proportion of its 
references through "Philosophy": $7.37$ million of the $7.4$ million references
are funneled through "Philosophy."
Second on the list of highest-ranking articles by traversal funnels is 
"Presentation" with only $30$ thousand paths. Similarly abstract 
ideas also rank highly such as "Tree of life" (30 thousand), 
"Reality" (13 thousand), "Jurisdiction" (3 thousand) also appear of the list.

>>>>

\title{\protect
Connecting Every Bit of Knowledge: \\
Wikipedia's First Link Network Unravelled
}

@@

"Philosophy" and other philosophical concepts with many traversal visits
are not among the highest-ranking articles by indegree.
"Philosophy" has an indegree of 581, with direct first links from articles about Philosophers and areas of Philosophy: "Existentialism and Humanism," "Predeterminism," "Synoptic Philosophy,"Qualia," "Dorothy Emmet," and "Christopher W. Morris."
While many articles accumulate at "Philosophy" (see traversal visits discussion), 
the accumulation is not the 
result of many articles directly referencing "Philosophy." 
Instead, the accumulation of first links, as we argue in our 
discussion of traversal visits, flows towards Philosophy as the 
ultimate anchor when genralizing from specific articles towards broader notions.
\begin{figure*}[tp!]
  \centering	
  \includegraphics[width=\columnwidth]{graphics/ndegree_loglog.png}
  \caption{
    \textbf{FLN Degree Distribution}
  }
  We construct a log-log plot fit our results to a linear model. The result is a 
  an excellent fit with r = -0.98, yielding a power law exponent of -0.79. 
  The distribution appears to be scale free with most articles having fewer than 9 
  direct first links, while few hold most.
  \label{fig:degree distribution}
\end{figure*}
The FLN's indegree exhibits a scale-free distribution where a few articles 
receive a most direct first references, while most articles receive few or none.
The average indegree for all 11 million articles is 3.6 direct first links with a standard deviation of 89.5 links.
Only 4,826 articles have more than 100 direct first links and $75\%$ of articles
have fewer than 9. 
When fit to a linear model on a log-log scale (log(rank) versus log(indegree)), 
the model's r value is $-0.98$ suggesting a strong log-log linear fit 
with a power law exponent of $-0.79$.

@@

As a distribution, the number traversal visits per article appears to be scale-free. The majority of articles have fewer than 30 traversal visits, while few 
have 5 order of magnitude more traversal visits. 
Specifically, $99.76\%$ of articles have fewer than $100$ traversal visits; nearly $80\%$ have none. 
Meanwhile, the highest ranking 30 articles have an extremely disproportionate number of traversal visits.

@@

\begin{figure*}[tp!]
  \centering	
  \includegraphics[width=\columnwidth]{graphics/synonyms.png}
  \caption{
    \textbf{Number of Synonyms for the highest-ranking articles}
    We use WordNet (the largest lexical database for the English language)
    to identify the number of synonyms for an article's title. We compare 
    the distribution of synonyms of the highest-ranking 100 articles
    against all other Wikipedia articles. As the distribution demonstrates
    the highest ranking concepts have more synonyms than other articles.
  }
  \label{fig:synonyms}
\end{figure*}

One means to measure the specificity of an article is to identify the number of synonyms available for a word or topic. The reasoning here is that a  broader topic likely has many more synonyms relative to a specific, concrete topic. Banana has fewer synonyms than botanical for example. To quantify the observed flow, we measured the number of synonyms each title contained in WordNet---the largest lexical database of the English language 
\cite{wordnet}. 
We find the highest ranking 100 articles have on average 5 more synonyms than the typical article; a difference of 2.5 synonyms if we compare the median number of synonyms in each group
(see figure~\ref{fig:synonyms})
. 
As suggested by the median, many articles have no synonyms as we might expect, because titles with more than one word are not likely to appear in a thesaurus. 
Since many articles have no synonyms, we also compared the number of synonyms in the highest-ranking versus typical article, this time excluding all articles without at least 1 synonym. 
We still find the highest ranking 100 articles with an average of 9.0 (median of 7.0) synonyms whereas the remaining articles on average 5.8 synonyms (median of 7.0), even with the exclusion of all articles without any synonyms.
The quantifiable difference in synonyms corroborates the flow of links from concrete, specific articles to broader disciplines or fundamental notions.

@@

\begin{figure*}[tp!]
  \centering	
  \includegraphics[width=\textwidth]{graphics/2_cycles.pdf}
  \caption{
    \textbf{highest ranking 2-cycles}
    We identify pairs of articles whose first links point to one another, forming
    a 2-cycle. We then rank each pair of articles by the total number of 
    traversal visits to gauge the most referenced groups of two articles linked
    to each other. We find are synonyms representing the same or nearly the 
    same concepts as opposed to distinct ideas.
  }
  \label{fig:2-cycles}
\end{figure*}

@@

\begin{figure*}[tp!]
  \centering	
  \includegraphics[width=\textwidth]{graphics/3_cycles.pdf}
  \caption{
    \textbf{highest ranking 3-cycles}
    We identify groups of three articles whose links form a 3-cycle with the 
    last article in the group linking back to the first article.
    We rank each 3-cycle by the total number of traversal visits to gauge
    the most referenced group of three linked articles.
  }
  \label{fig:3-cycles}
\end{figure*}

@@

Ranking articles by the number of traversal funnels we find 
"Philosophy" as by far the highest-ranking article with 
$7.37$ million paths
(see figure~\ref{fig:Funnels}).
\begin{figure*}[tp!]
  \centering	
  \includegraphics[width=\textwidth]{graphics/funnels.pdf}
  \caption{
    \textbf{Funnels}
    We represent the highest-ranking articles by the number of traversal 
    funnels to gauge the influence each article exerts in shaping the 
    structure of the FLN. We find "Philosophy" exerts an overwhelming proportion
    of the influence with other abstract notions and topical concepts ranking
    next.
  }
  \label{fig:Funnels}
\end{figure*}
Of any article, the number of traversal funnels Philosophy holds exceeds 
all others by at least two orders of magnitude.
The "Philosophy" cycle which contains "Existence," "Awareness," "Reality," 
and similar articles accumulates the overwhelming proportion of its 
references through "Philosophy": $7.37$ million of the $7.4$ million references
are funneled through "Philosophy."
Second on the list of highest-ranking articles by traversal funnels is 
"Presentation" with only $30$ thousand paths. Similarly abstract 
ideas also rank highly such as "Tree of life" (30 thousand), 
"Reality" (13 thousand), "Jurisdiction" (3 thousand) also appear of the list.
